---
title: "Final project report"
author: "Siting Lin, Nitin Jenson, Brittnie Yi"
output: html_notebook
---

### Research Question:

How do external, geographic factors, and pre-existing medical conditions affect the number of cases and death rates from the Coronavirus infection?

### Primary data:

This dataset is taken from https://www.worldometers.info/coronavirus/
The .CSV file for this dataset is called worldomemtersData.csv in the Github Repo.

### Secondary data:

This dataset is taken from https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3551445/
The .CSV file for this dataset is called ICUnumbers.csv in the Github Repo.

This dataset is taken from https://www.kaggle.com/folaraz/world-countries-and-continents-details
The .CSV file for this dataset is called countries and continents.csv in the Github Repo.


```{r}
# Load all packages needed
library(tidyverse)
library(ggplot2)
library(DataComputing)
library(rworldmap)
library(mosaic)
library(readr)
```

```{r}
# This step loads the primary and secondary datasets. Since the primary data will be updated everday, we chose to use data from the day that we finished the preliminary EDA.

file_name <- file.choose() 
Worldmeters_data <- 
  data.table::fread(file_name)  
Worldmeters_data

file_name2 <- file.choose() 
ICUnumbers <- 
  data.table::fread(file_name2)  
ICUnumbers

file_name3 <- file.choose() 
Continents <- 
  data.table::fread(file_name3)  
Continents
```

```{r}
# Inspection of imported data before analysis
# Using the function str() helps us see the structure of each dataset, such as its variables and variable types. 

str(Worldmeters_data)
str(ICUnumbers)
str(Continents)
```

The World Meters dataset contains 212 observations, 12 variables, and most of these variables are of 'chr' type besides the variable Deaths/1Mpop. The ICU numbers dataset contains 14 observations and 4 variables of different types. The Continents dataset contains 251 observations, which are all the different countries, and 27 variables.

```{r}
# This step cleans the data that we are going to use for analysis

ICUnew <-
  ICUnumbers %>%
  rename( Healthcare_cost = 'Per Capita Healthcare Cost**' ) %>% # Rename the variable for easy use
  mutate( Healthcare_cost = gsub(pattern = "[$]", replacement = " ", Healthcare_cost )) %>% # Edit all regular expressions
  rename( 'ICUbeds' = 'ICU beds per 100,000 people' )# For easy use, we abbreviated the variable's name
ICUnew

# Rename the variables in the data frame for ease and clarity, and select variables for analysis

Worldnew <- 
  Worldmeters_data %>%
  rename( Total_Cases = `Total Cases` ) %>%
  rename( Total_Death = `Total Deaths` ) %>%
  rename( Total_Recovered = 'Total Recovered' ) %>%
  rename( Total_Tests = `Total Tests` ) %>%
  select( Country, Total_Cases, Total_Death,Total_Recovered,Total_Tests)
Worldnew

# Repeat the renaming

Continent_new<-
  Continents %>%
  select( name, Continent ) %>%
  rename( Country = name )
Continent_new

```
```{r}
# Convert all characters to integers or numbers that can be used in analysis, and join ICU dataset and World Meters dataset

Combine_data <-
  ICUnew %>%
  inner_join( Worldnew, by = "Country" ) %>%
  filter( !Country == "Zambia" ) %>% # Zambia's data is not available in this set, so we deleted it
  mutate( Total_Cases = parse_number(Total_Cases) ) %>%
  mutate( Total_Death = parse_number(Total_Death) ) %>%
  mutate( Total_Recovered = parse_number(Total_Recovered) ) %>%
  mutate( Total_Tests = parse_number(Total_Tests) ) %>%
  mutate( Healthcare_cost = parse_number(Healthcare_cost) )
  
Combine_data

```


# The relationship between external factors and pre_existing medical conditions

We are going to explore the relationship between external factors, such as healthcare cost, pre-existing medical conditions, using ICU numbers of 100000 people, confirmed cases, and death rate.

### Step 1: The relationship between ICU numbers, death rate, and pre_existing medical conditions

The reason that we are using these data sets is because more developed countries with larger populations are able to provide more accurate data about ICU beds, health care costs, and data about confirmed coronavirus cases. There are a few possibilities about how humans can handle COVID 19, such as having a strong immune system and recovering from the virus in isolation, or be tranfered to the ICU that can provide better medical services and care. Therefore, we hypothesize that the numbers of ICU beds avaliable and variables such as death rate in different countries may have a relationship,and we are going to explore that using the data.

```{r}
# Add the new variable called death rate

ICU_Death_Rela <-
  Combine_data %>%
  select( ICUbeds, Total_Death, Total_Cases, Country ) %>%
  mutate( Deathrate = Total_Death / Total_Cases )

# Create a visualization of the relationship between death rate and ICU bed numbers in specific countries

ggplot( data = ICU_Death_Rela, aes(x = ICUbeds, y = Deathrate) ) +
  geom_point( aes(color = Country) ) +
  geom_smooth( method = lm ) +
  xlab("ICU beds")

```

There is a small number of data points due to the ICU numbers data set, which could be the reason why there is no obvious relationship between death rate and ICU numbers average. We feel that the ICU numbers data set contains accurate records, so in this analysis, we are trading a large number of data points for accuracy of the actual data. The relationship shown in the graph is positive, indicating that our hypothesis was false and that there is little relationship between ICU beds and the death rate. However, if we try to exclude some outliers, that is, to constrain the range of ICU beds from just 6 to 10, we might be able to see some form of a relationship.

```{r}
ggplot( data = ICU_Death_Rela, aes(x = ICUbeds, y = Deathrate) ) +
  geom_point( aes(color = Country) ) +
  geom_line() + 
  geom_smooth( method = lm ) +
  xlim( 6, 10) +
  xlab("ICU beds")
```

When we tighten the range of ICU bed numbers, there is a more obvious decresing relationship between death rate and ICU beds. However, this does not tell the whole story. Since we excluded multiple countries in this graph, coming to the conclusion of ICU bed number being related to deathrate has little accuracy and is not reliable. This showed us that there are lots of factors that were not considered that would be affecting our results, such as when these countries took action in quarantine orders. Since the data we used is already outdated, it also proves how important updating data and maintaining a database is when doing analysis in real life situations. 

When looking at the previous visualization with less of a relationship, some of the outlier countries had a much different process in handling the virus than countries that did follow the trend that we 'forced' in the next visualization. Countries such as Spain felt the impact of the virus before the United States, and that could be why they have such a high death rate, even with more ICU bed numbers. 

### Step 2: The relationship between average healthcare cost and total comfirmed cases.

Since the tests that comfirmed the infection of CoronaVirus are expensive for most population, the countries that have higher healthcare cost may afford more tests for their population. We assume there is a relationship between total tests and healthcare cost in each countries.

```{r}
Health_Test_Rela<-
  Combine_data%>%
  select(Country, Healthcare_cost, Total_Tests)%>%
  mutate(Total_Tests_rate = Total_Tests/100)# for better illustrate in graph, we changed the variable of total tests to the tests rate of 100 people
ggplot(data = Health_Test_Rela, aes(x = Total_Tests_rate, y = Healthcare_cost))+
  geom_point(aes(color = Country))
```

According to the graph shown above, there is no obivious relationship between total tests rate and healthcare cost.

# The relationship between geographic factor and numbers of cases comfirmed


For futher exploring the relationship between geographic factor(continents) and total cases comfirmed, we are going to make a graph. 

```{r}
Con_Case_Rela<-
  Continent_new%>%
  inner_join(Worldnew)%>%
  mutate(Total_Cases = parse_number(Total_Cases))%>%
  mutate(Total_Death = parse_number(Total_Death))%>%
  filter(!Total_Death =="NA")%>%
  group_by(Continent)%>%
  summarise(Total_Cases_conti = sum(Total_Cases),
            Total_Death_conti = sum(Total_Death))%>%
  mutate(Deathrate_conti = Total_Death_conti/Total_Cases_conti)
ggplot(data = Con_Case_Rela, aes(x = Deathrate_conti, y = Total_Cases_conti))+
  geom_point(aes(color = Continent))
Con_Case_Rela
```


According to the graph above, we can know that there is a largeer population in Asia got comfirmed. Other total cases in each Continent are similar. However, there is a outlier which is Europe showing a really high death rate.
